Novel lattice Boltzmann method for simulation of strongly shear thinning viscoelastic fluids

### 2024-06-22 NovelLBM.tex
fix whitespace

Our next example is the RT-DC geometry in which cells from a reservoir are squeezed through a narrow observation channel \cite{Toepfner.2018, Fregin:2019gx} with a square cross-section of %\qtyproduct{40x40}{\micro\meter}.
%$\SI{40}{\micro\meter} \times \SI{40}{\micro\meter}$.%Correctly typeset version, that compiles for Stephan.
For this simulation we use periodic boundary conditions and a constant body force corresponding to a pressure gradient of \SI{1e7}{\pascal\per\meter}.
In figure~\ref{fig:RT-DC}, we show the velocity profile, the von Mises stress and the shear component of the stress which is the relevant component if a local shear viscosity is to be defined.
There are three qualitative differences compared to a purely viscous fluid.
Directly after the inlet we find a region of low stress.
Due to the fast advection, the polymers, despite being inside the constriction, have not reacted to the increased fluid stress yet.
Similarly, directly after the constriction, the high stress inside the observation channel persists over a certain distance that is needed for the polymer stress to decay. If one would divide the stress in these two regions by the strain rate in an attempt to obtain a viscosity one would find a very low and a very high viscosity respectively. These can be outside the range of viscosities, that can be obtained through rheological measurements with the fluid. This effect is even more pronounced along the walls directly after the outlet. Here the viscosities obtained in this manner become negative. Due to the strain rate approaching $0$ at several locations within the simulation volume, this calculation often diverges. The clamped result can be seen in figure~\ref{fig:Nozzle/RT-DC_eta}. These discrepancies, especially the ones at the walls, show the breakdown of the viscosity interpretation in viscoelastic cases and signify the regions of interesting behavior. This also shows why viscoelastic simulations cannot always be replaced by purely viscous models and should be considered despite the high computational cost.

>>>>

Our next example is the RT-DC geometry in which cells from a reservoir are squeezed through a narrow observation channel \cite{Toepfner.2018, Fregin:2019gx} with a square cross-section of \qtyproduct{40x40}{\micro\meter}.
%$\SI{40}{\micro\meter} \times \SI{40}{\micro\meter}$.%Correctly typeset version, that compiles for Stephan.
For this simulation we use periodic boundary conditions and a constant body force corresponding to a pressure gradient of \SI{1e7}{\pascal\per\meter}.
In figure~\ref{fig:RT-DC}, we show the velocity profile, the von Mises stress and the shear component of the stress which is the relevant component if a local shear viscosity is to be defined.
There are three qualitative differences compared to a purely viscous fluid.
Directly after the inlet we find a region of low stress.
Due to the fast advection, the polymers, despite being inside the constriction, have not reacted to the increased fluid stress yet.
Similarly, directly after the constriction, the high stress inside the observation channel persists over a certain distance that is needed for the polymer stress to decay. If one would divide the stress in these two regions by the strain rate in an attempt to obtain a viscosity one would find a very low and a very high viscosity respectively. These can be outside the range of viscosities, that can be obtained through rheological measurements with the fluid. This effect is even more pronounced along the walls directly after the outlet. Here the viscosities obtained in this manner become negative. Due to the strain rate approaching $0$ at several locations within the simulation volume, this calculation often diverges. The clamped result can be seen in figure~\ref{fig:Nozzle/RT-DC_eta}. These discrepancies, especially the ones at the walls, show the breakdown of the viscosity interpretation in viscoelastic cases and signify the regions of interesting behavior. This also shows why viscoelastic simulations cannot always be replaced by purely viscous models and should be considered despite the high computational cost.
